---
title: "Colonization"
output: html_notebook
---

This first command are used to clean-up the memory and start a new session
```{r}
rm(list = ls())
dev.off()
```

Loading the required packages and importing the data
```{r}
library(readxl)
library(stats) #statistical package
library(MASS) #boxcox transformation

library(multcompView)
library(lsmeans)
library(multcomp)
library(ggplot2)
library(psych)
library(plyr)
library(tidyverse)
library(car)
library(ggpubr)
library(sandwich)

library(dplyr) #summarized and grouped tables
library(magrittr)

library(lubridate) #required to not getcrazy with datas
library(scales)

#library(QTLRel)
#importing the dataframe
df <- read_excel("/Volumes/SGlab_Geier/GitHub/Colonization.xlsx")
View(df) 
```

Converting column types
```{r}
df$day<-ymd(df$day)
df$day_number<-as.factor(df$day_number)
df$part<- as.factor(df$part)
df$value <- as.numeric(as.character(df$value))
```

Normality test on each dataset
```{r}
shapiro.test(df$value[df$thesis == "1"])
shapiro.test(df$value[df$thesis == "2"])
shapiro.test(df$value[df$thesis == "3"])
shapiro.test(df$value[df$thesis == "4"])
shapiro.test(df$value[df$thesis == "5"])
shapiro.test(df$value[df$thesis == "6"])
shapiro.test(df$value[df$thesis == "7"])
shapiro.test(df$value[df$thesis == "8"])
shapiro.test(df$value[df$thesis == "9"])
shapiro.test(df$value[df$thesis == "10"])
shapiro.test(df$value[df$thesis == "11"])
shapiro.test(df$value[df$thesis == "12"])
shapiro.test(df$value[df$thesis == "13"])
shapiro.test(df$value[df$thesis == "14"])
shapiro.test(df$value[df$thesis == "15"])
shapiro.test(df$value[df$thesis == "16"])
shapiro.test(df$value[df$thesis == "17"])
shapiro.test(df$value[df$thesis == "18"])
shapiro.test(df$value[df$thesis == "19"])
shapiro.test(df$value[df$thesis == "20"])
shapiro.test(df$value[df$thesis == "21"])
shapiro.test(df$value[df$thesis == "22"])
shapiro.test(df$value[df$thesis == "23"])
shapiro.test(df$value[df$thesis == "24"])
shapiro.test(df$value[df$thesis == "25"])
shapiro.test(df$value[df$thesis == "26"])
shapiro.test(df$value[df$thesis == "27"])
shapiro.test(df$value[df$thesis == "28"])

```

The data are not normal. Use BocCox transformation to put them back in line
```{r}
#generate a new dataframe to work on
ds <- as.data.frame(df)

#generate the model on the value dependig upon the sampling data (day) and the compartment (part) variables
my_model<-lm(value ~ day * part, data=ds)
bc1 <- boxcox(my_model, lambda= seq(-3,3), ploit=TRUE)
#find the bestlambda value for thetransformation
bes.lam<-bc1$x[which(bc1$y==max(bc1$y))]

#generate the new model and plot it to be sure that it works
fullmodel<-lm((ds$value)^bes.lam~day*part, data=ds)
plot(fullmodel)

#generate the new column with the normalized value in the original df
df["norm"]<-df$value**bes.lam
```

Normality test on each dataset post transformation
```{r}
shapiro.test(df$value[df$thesis == "1"])
shapiro.test(df$value[df$thesis == "2"])
shapiro.test(df$value[df$thesis == "3"])
shapiro.test(df$value[df$thesis == "4"])
shapiro.test(df$value[df$thesis == "5"])
shapiro.test(df$value[df$thesis == "6"])
shapiro.test(df$value[df$thesis == "7"])
shapiro.test(df$value[df$thesis == "8"])
shapiro.test(df$value[df$thesis == "9"])
shapiro.test(df$value[df$thesis == "10"])
shapiro.test(df$value[df$thesis == "11"])
shapiro.test(df$value[df$thesis == "12"])
shapiro.test(df$value[df$thesis == "13"])
shapiro.test(df$value[df$thesis == "14"])
shapiro.test(df$value[df$thesis == "15"])
shapiro.test(df$value[df$thesis == "16"])
shapiro.test(df$value[df$thesis == "17"])
shapiro.test(df$value[df$thesis == "18"])
shapiro.test(df$value[df$thesis == "19"])
shapiro.test(df$value[df$thesis == "20"])
shapiro.test(df$value[df$thesis == "21"])
shapiro.test(df$value[df$thesis == "22"])
shapiro.test(df$value[df$thesis == "23"])
shapiro.test(df$value[df$thesis == "24"])
shapiro.test(df$value[df$thesis == "25"])
shapiro.test(df$value[df$thesis == "26"])
shapiro.test(df$value[df$thesis == "27"])
shapiro.test(df$value[df$thesis == "28"])

```

Homoschedasticy test
```{r}
#we can decide wether to use Barlett's test or Levene's test. Barlett is extremely conservative, but I like Levene's better 
leveneTest(norm~part, data=df)
leveneTest(norm~day, data=df)

```

Now that the datas are normalized we proceed with the statistical analysis
```{r}
mod<-aov(norm~day*part, data=df)
Anova(mod, type="III")
summary(mod)
marginal= lsmeans(mod, ~day*part)
tuk.cld<-cld(marginal, alpha=.05, Letters=letters)

```

Graphs prerequisites
```{r}
#check that all the column are still in the right format
str(df)

#add the log column, that we will use to visualize the data
df["Log"]<-log(df$value)

#add the DPI column
df["DPI"]<- "15DPI"
df[df$day_number == 2 , "DPI"] <- "18 DPI"
df[df$day_number == 3 , "DPI"] <- "22 DPI"
df[df$day_number == 4 , "DPI"] <- "29 DPI"
df[df$day_number == 5 , "DPI"] <- "36 DPI"
df[df$day_number == 6 , "DPI"] <- "43 DPI"
df[df$day_number == 7 , "DPI"] <- "57 DPI"

#create a df (T) based on df (df) grouped by thesis (part*day). The result is visually no different than the original
TS<-df%>%group_by(thesis)

#create a summarized data table (SUM)
SUM<-TS%>%summarise(day=first(day), day_number=first(day_number), DPI=first(DPI), part=first(part), Mean=mean(Log), Sem=sd(Log)/sqrt(length(Log)))
```

Graphs by compartment (part)
```{r}
#create column for dpi as factor (cond) 
df$cond <- as.factor(df$DPI)

#Generate the part subsets
sA <- subset(df, part=="A") #apexes
sPa <- subset(df, part=="Pa") #upper leaves
sPb <- subset(df, part=="Pb") #lower leaves
sR <- subset(df, part=="R") #roots

#define the model for each subset (value ~ dpi variable, set in the condidtion)
#ANOVA on the model
mA<-lm(Log~cond, data=sA)
mPa<-lm(Log~cond, data=sPa)
mPb<-lm(Log~cond, data=sPb)
mR<-lm(Log~cond, data=sR)

Anova(mA, type="III")
Anova(mPa, type="III")
Anova(mPb, type="III")
Anova(mR, type="III")

summary(mA)
summary(mPa)
summary(mPb)
summary(mR)


#least square means of each model depending on the sampling day
marginalA= lsmeans(mA, ~cond)
marginalPa= lsmeans(mPa, ~cond)
marginalPb= lsmeans(mPb, ~cond)
marginalR= lsmeans(mR, ~cond)

#show structure
str(marginalA) 
str(marginalPa)
str(marginalPb)
str(marginalR)

#compact letters display
#returns the statistical differences as letters
tukA<-cld(marginalA, alpha=.05, Letters=letters)
tukPa<-cld(marginalPa, alpha=.05, Letters=letters)
tukPb<-cld(marginalPb, alpha=.05, Letters=letters)
tukR<-cld(marginalR, alpha=.05, Letters=letters)

#Order tuk results by cond (so by dpi) 
tukA<-tukA[order(tukA$cond), ]
tukPa<-tukPa[order(tukPa$cond), ]
tukPb<-tukPb[order(tukPb$cond), ]
tukR<-tukR[order(tukR$cond), ]

#subsets of the summarized table (already ordered)
ssA<-subset(SUM, part=="A") 
ssPa<-subset(SUM, part=="Pa")
ssPb<-subset(SUM, part=="Pb")
ssR<-subset(SUM, part=="R")

#now that both are ordered, move the column group in the subset 
ssA["Lab"]<-tukA$.group
ssPa["Lab"]<-tukPa$.group
ssPb["Lab"]<-tukPb$.group
ssR["Lab"]<-tukR$.group

#Production of the graphs for each plant compartment in time
#For the apex
ggplot(ssA, aes(x=day, y=Mean, label=rownames(Lab))) +
  geom_errorbar(aes(x=day, ymin=Mean-Sem, ymax=Mean+Sem), alpha=.9) +
  geom_bar(position="dodge", stat="identity", color= "black", fill="light gray") +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x=element_text(angle = 0)) +
  scale_x_date(labels = date_format("%m/%d"), breaks = ssA$day)+
  scale_y_continuous(breaks = seq(0, 18, 2))+
  theme(legend.position = "none") +
  geom_text(label=ssA$Lab, hjust = 0.5, nudge_y = 2.5) +
  labs(x="DPI", y="Log(GU/ng)", title= "Apex")

#For the upper leaves
ggplot(ssPa, aes(x=day, y=Mean, label=rownames(Lab))) +
  geom_errorbar(aes(x=day, ymin=Mean-Sem, ymax=Mean+Sem), alpha=.9) +
  geom_bar(position="dodge", stat="identity", color= "black", fill="light gray") +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x=element_text(angle = 0)) +
  scale_x_date(labels = date_format("%m/%d"), breaks = ssPa$day)+
  scale_y_continuous(breaks = seq(0, 18, 2))+
  theme(legend.position = "none") +
  geom_text(label=ssPa$Lab, hjust = 0.5, nudge_y = 2.5) +
  labs(x="DPI", y="Log(GU/ng)", title= "Upper Leaves")

#For the lower leaves
ggplot(ssPb, aes(x=day, y=Mean, label=rownames(Lab))) +
  geom_errorbar(aes(x=day, ymin=Mean-Sem, ymax=Mean+Sem), alpha=.9) +
  geom_bar(position="dodge", stat="identity", color= "black", fill="light gray") +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x=element_text(angle = 0)) +
  scale_x_date(labels = date_format("%m/%d"), breaks = ssPb$day)+
  scale_y_continuous(breaks = seq(0, 18, 2))+
  theme(legend.position = "none") +
  geom_text(label=ssPb$Lab, hjust = 0.5, nudge_y = 2.5) +
  labs(x="DPI", y="Log(GU/ng)", title= "Lower Leaves")

#For the roots
ggplot(ssR, aes(x=day, y=Mean, label=rownames(Lab))) +
  geom_errorbar(aes(x=day, ymin=Mean-Sem, ymax=Mean+Sem), alpha=.9) +
  geom_bar(position="dodge", stat="identity", color= "black", fill="light gray") +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x=element_text(angle = 0)) +
  scale_x_date(labels = date_format("%m/%d"), breaks = ssR$day)+
  scale_y_continuous(breaks = seq(0, 18, 2))+
  theme(legend.position = "none") +
  geom_text(label=ssR$Lab, hjust = 0.5, nudge_y = 2.5) +
  labs(x="DPI", y="Log(GU/ng)", title= "Roots")
```

Graphs by sampling (dpi)
```{r}
#create the column for plant compartment as a factor (cond)
SUM$cond=as.factor(SUM$part)

#Generate the sampling date subsets
st1<-subset(df, day_number=="1")
st2<-subset(df, day_number=="2")
st3<-subset(df, day_number=="3")
st4<-subset(df, day_number=="4")
st5<-subset(df, day_number=="5")
st6<-subset(df, day_number=="6")
st7<-subset(df, day_number=="7")

st1$cond<-st1$part
st2$cond<-st2$part
st3$cond<-st3$part
st4$cond<-st4$part
st5$cond<-st5$part
st6$cond<-st6$part
st7$cond<-st7$part

#define the model for each subset (value ~ dpi variable, set in the condidtion)
#ANOVA on the model
mod1<-lm(Log~part, data=st1)
mod2<-lm(Log~part, data=st2)
mod3<-lm(Log~part, data=st3)
mod4<-lm(Log~part, data=st4)
mod5<-lm(Log~part, data=st5)
mod6<-lm(Log~part, data=st6)
mod7<-lm(Log~part, data=st7)

Anova(mod1, type="III")
Anova(mod2, type="III")
Anova(mod3, type="III")
Anova(mod4, type="III")
Anova(mod5, type="III")
Anova(mod6, type="III")
Anova(mod7, type="III")

summary(mod1)
summary(mod2)
summary(mod3)
summary(mod4)
summary(mod5)
summary(mod6)
summary(mod7)

#least square means of each model depending on the plant compartment
marginal1= lsmeans(mod1, ~part)
marginal2= lsmeans(mod2, ~part)
marginal3= lsmeans(mod3, ~part)
marginal4= lsmeans(mod4, ~part)
marginal5= lsmeans(mod5, ~part)
marginal6= lsmeans(mod6, ~part)
marginal7= lsmeans(mod7, ~part)

#compact letters display
#returns the statistical differences as letters
tuk1<-cld(marginal1, alpha=.05, Letters=letters)
tuk2<-cld(marginal2, alpha=.05, Letters=letters)
tuk3<-cld(marginal3, alpha=.05, Letters=letters)
tuk4<-cld(marginal4, alpha=.05, Letters=letters)
tuk5<-cld(marginal5, alpha=.05, Letters=letters)
tuk6<-cld(marginal6, alpha=.05, Letters=letters)
tuk7<-cld(marginal7, alpha=.05, Letters=letters)

#Order tuk results by compartment
tuk1<-tuk1[order(tuk1$part), ]
tuk2<-tuk2[order(tuk2$part), ]
tuk3<-tuk3[order(tuk3$part), ]
tuk4<-tuk4[order(tuk4$part), ]
tuk5<-tuk5[order(tuk5$part), ]
tuk6<-tuk6[order(tuk6$part), ]
tuk7<-tuk7[order(tuk7$part), ]

#subsets of the summarized table (already ordered)
ss1<-subset(SUM, day_number=="1")
ss2<-subset(SUM, day_number=="2")
ss3<-subset(SUM, day_number=="3")
ss4<-subset(SUM, day_number=="4")
ss5<-subset(SUM, day_number=="5")
ss6<-subset(SUM, day_number=="6")
ss7<-subset(SUM, day_number=="7")

#now that both are ordered, move the column group in the subset 
ss1["Lab"]<-tuk1$.group
ss2["Lab"]<-tuk2$.group
ss3["Lab"]<-tuk3$.group
ss4["Lab"]<-tuk4$.group
ss5["Lab"]<-tuk5$.group
ss6["Lab"]<-tuk6$.group
ss7["Lab"]<-tuk7$.group

#Production of the graphs for each sampling date
#15DPI
ggplot(ss1, aes(x=part, y=Mean, fill=cond, label=rownames(Lab))) +
  geom_errorbar(aes(x=part, ymin=Mean-Sem, ymax=Mean+Sem), position = "dodge", alpha=.7) +
  geom_bar(position="dodge", stat="identity") +
  labs(x="Plant Compartment", y="Log (GU/ng)", title= "15 DPI") +
  geom_text(label=ss1$Lab, hjust = 0.5, nudge_y = 2.0) +
  scale_fill_grey(start=0.8, end=0.2)+
  scale_x_discrete(labels = c("Apex", "Upper Leaves", "Lower Leaves", "Roots"))+
  theme_light() +
  theme(legend.position = "none") +
  ylim(0,19)

#18DPI
ggplot(ss2, aes(x=part, y=Mean, fill=cond, label=rownames(Lab))) +
  geom_errorbar(aes(x=part, ymin=Mean-Sem, ymax=Mean+Sem), position = "dodge", alpha=.7) +
  geom_bar(position="dodge", stat="identity") +
  labs(x="Plant Compartment", y="Log (GU/ng)", title= "18 DPI") +
  geom_text(label=ss2$Lab, hjust = 0.5, nudge_y = 2.0) +
  scale_fill_grey(start=0.8, end=0.2)+
  scale_x_discrete(labels = c("Apex", "Upper Leaves", "Lower Leaves", "Roots"))+
  theme_light() +
  theme(legend.position = "none")+
  ylim(0,19)

#22DPI
ggplot(ss3, aes(x=part, y=Mean, fill=cond, label=rownames(Lab))) +
  geom_errorbar(aes(x=part, ymin=Mean-Sem, ymax=Mean+Sem), position = "dodge", alpha=.7) +
  geom_bar(position="dodge", stat="identity") +
  labs(x="Plant Compartment", y="Log (GU/ng)", title= "22 DPI") +
  geom_text(label=ss3$Lab, hjust = 0.5, nudge_y = 2.0) +
  scale_fill_grey(start=0.8, end=0.2)+
  scale_x_discrete(labels = c("Apex", "Upper Leaves", "Lower Leaves", "Roots"))+
  theme_light() +
  theme(legend.position = "none") +
  ylim(0,19)

#29DPI
ggplot(ss4, aes(x=part, y=Mean, fill=cond, label=rownames(Lab))) +
  geom_errorbar(aes(x=part, ymin=Mean-Sem, ymax=Mean+Sem), position = "dodge", alpha=.7) +
  geom_bar(position="dodge", stat="identity") +
  labs(x="Plant Compartment", y="Log (GU/ng)", title= "29 DPI") +
  geom_text(label=ss4$Lab, hjust = 0.5, nudge_y = 2.0) +
  scale_fill_grey(start=0.8, end=0.2)+
  scale_x_discrete(labels = c("Apex", "Upper Leaves", "Lower Leaves", "Roots"))+
  theme_light() +
  theme(legend.position = "none") +
  ylim(0,19)

#36DPI
ggplot(ss5, aes(x=part, y=Mean, fill=cond, label=rownames(Lab))) +
  geom_errorbar(aes(x=part, ymin=Mean-Sem, ymax=Mean+Sem), position = "dodge", alpha=.7) +
  geom_bar(position="dodge", stat="identity") +
  labs(x="Plant Compartment", y="Log (GU/ng)", title= "36 DPI") +
  geom_text(label=ss5$Lab, hjust = 0.5, nudge_y = 2.0) +
  scale_fill_grey(start=0.8, end=0.2)+
  scale_x_discrete(labels = c("Apex", "Upper Leaves", "Lower Leaves", "Roots"))+
  theme_light() +
  theme(legend.position = "none") +
  ylim(0,19)

#43DPI
ggplot(ss6, aes(x=part, y=Mean, fill=cond, label=rownames(Lab))) +
  geom_errorbar(aes(x=part, ymin=Mean-Sem, ymax=Mean+Sem), position = "dodge", alpha=.7) +
  geom_bar(position="dodge", stat="identity") +
  labs(x="Plant Compartment", y="Log (GU/ng)", title= "43 DPI") +
  geom_text(label=ss6$Lab, hjust = 0.5, nudge_y = 2.0) +
  scale_fill_grey(start=0.8, end=0.2)+
  scale_x_discrete(labels = c("Apex", "Upper Leaves", "Lower Leaves", "Roots"))+
  theme_light() +
  theme(legend.position = "none") +
  ylim(0,19)

#57DPI
ggplot(ss7, aes(x=part, y=Mean, fill=cond, label=rownames(Lab))) +
  geom_errorbar(aes(x=part, ymin=Mean-Sem, ymax=Mean+Sem), position = "dodge", alpha=.7) +
  geom_bar(position="dodge", stat="identity") +
  labs(x="Plant Compartment", y="Log (GU/ng)", title= "57 DPI") +
  geom_text(label=ss7$Lab, hjust = 0.5, nudge_y = 2.0) +
  scale_fill_grey(start=0.8, end=0.2)+
  scale_x_discrete(labels = c("Apex", "Upper Leaves", "Lower Leaves", "Roots"))+
  theme_light() +
  theme(legend.position = "none") +
  ylim(0,19)


```